Learning by Teaching a Robot: the Case of Handwriting

### 2015-07-10 main.tex
Changed children's names + blurred face

Henry is five and an half, and has been diagnosed with visuo-constructive
deficits. He is under the care of an occupational therapist, and tries to
workaround his inability to draw letters in a consistent manner. Diego is six
and struggles at school with his poor handwriting and even poorer
self-confidence.

While Henry is lively and always quick at shifting his attention from one
activity to another, Diego is shy and poised. Two very different children,
facing however the same difficulty to write in a legible manner. And, hidden
beyond this impaired skill, psycho-social difficulties also arise: they
underperform at school, Henry has to go for follow-up visits every week,
they both live under the label ``requires special care''. This is a source of
anxiety for the children and for their parents alike.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{henry}
    \caption{\small Henry teaching Nao how to write numbers, with the help of an
    occupational therapist.}
    \label{fig:henry}
\end{figure}

@@

\begin{table}[ht!]
\centering
\caption{\small Field studies conducted within the project}
\label{studies}
\footnotesize
\begin{tabular}{@{}lp{4cm}p{2.2cm}p{1.8cm}l@{}}
\toprule
{\bf Study}     & {\bf Type}                                    & {\bf Duration}              & {\bf Children \#}                     & {\bf Ages} \\ \midrule
{\it ISG 1}     & Unstructured group interaction at school      & 16 min/group               & 4 $\times$ 8 children                 & 6-7        \\
{\it Florimont} & Individual/Pair interaction at school         & 11 min/group               & 7 (individual) + 7 $\times$ 2 (pairs) & 7-8        \\
{\it Diego}     & Case-study, spanning over 4 weeks              & 1.5 hours $\times$ 4 weeks & 1                                     & 5          \\
{\it Haut-Lac}  & Pair (with turn-taking) interaction at school & 26 min/group                            & 7 $\times$ 2                          & 5-6        \\
{\it ISG 2}     & Individual interaction at school              & 18 min                            & 6                                     & 5-6        \\
{\it Henry}     & Case-study, spanning over 4 weeks              & 1h $\times$ 4 weeks        & 1                                     & 5          \\ \bottomrule
\end{tabular}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Case Study 1: Diego}

@@

Diego is a six years old child with troubles in handwriting and a poor
self-esteem. Our goal was to figure out an environment in
order to make Diego sustaining engagement over four sessions of one hour, one
session per week. We decided to introduce an appealing scenario that justifyed to the child
the activity where a robot wants to learn handwriting. We used two Nao robots :
a blue one (called Mimi) and an orange one (called Clem). Mimi was away for a
mission,and the two robots had to communicate by mails. But they decided to do
it ``like humans", with handwritten messages. While Mimi was a good in
handwriting, Clem had strong difficulties and needed the help of Diego.\\

During the first session, Diego just teached single letters to the robot.
To chose a letter, he had to pick it in a set of magneted plastic letters, and to
show it to the robot. But the robot was not programmed to detect and recognize
it. A third person was entering the letter and was sending it to the
system with the computer. With following sessions we gradually increased the
difficulty : in the second session the child teached full words. To chose the
word, he had to assemble plastic letters to build it and to show it to the
robot. In the two last sessions, he teached long sentences. This time he just
had to say the sentence to the robot, and then the third person was using the
compurter to enter it. \\

We designed a special scenario for the last session : Mimi came back to the
laboratory. In front of Diego, Mimi said to Clem that it did not believe 
it realy improved its handwritting, and that the responses it recieved
was written by Diego. To prove that it actualy improved its writing,
it gave to Clem thirty minutes for a final training. After the training, it
asked to Clem to write some words. Finally, he admitted that Clem actually
improved its handwriting.

\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{diego}
    \caption{\small Diego working with Nao}
    \label{fig:diego}
\end{figure}

\subsubsection{Results}
In all sessions, Diego was focused on the activity during forty minutes up to
one hour. In total, he provided 154 letters as demonstration.\\

Diego accepted our scenario and seriously played the game. At the
beginning he seemed not confident and intimidated.  But quickly, he gain in
confidence to play with the robot. He even showed real affective bonds : he
sent an handwritten mail to it some days after the last session.\\

As a technical result, we do not know if Diego improved his handwriting but he
likely conviced himself he was good enough at writing to help someone else (the
robot in that case). To show to Diego how much he improved the robot's
handwriting, we printed the letter that the robot wrote during the third session
before and after his help (Fig.~\ref{fig:stimuli}).

\begin{figure}
    \centering
    \subfigure[Initial letter, generated by the robot]{
        \includegraphics[height=6cm]{diego-initial-letter}
    }
    \subfigure[Final letter, after training with Diego]{
        \includegraphics[height=6cm]{diego-final-letter}
    }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Case Study 2: Henry}

\subsubsection{Context, Study Design}
Henry is a five and half years old child. It is followed by an occupational
therapist. According to the therapist, he presents visuo-constructive deficits.
From our perspective, the main effect of this trouble during handwriting
activity was that Henry
did not have any mental template to draw the shape of a letter. So he was 
apparently performing random attempts and then was comparing with the provided template.
What is more, Henry is strongly careless : he rarely payed attention to
advices, even to what he was doing when he was currently drawing, and he was
quickly shifting his attention from one
activity to another.
Hence, the goal of this study is to face those two challenging issues
(visuo-constructive deficits and inattention) in order to keep Henry focused
on the activity during forty-minutes interactions, and to make the robot
evidently learning from his demonstrations.\\

This experiment took place in the therapist's office. It was divided in four
sessions as well. This time, we assumed that a scenario like the one we used for Diego was
no longer relevant with Henry. We just introduced the robot and quickly
said that it was seeking help to train for a robot handwriting contest.\\

Before the experiment, Henry was working on writing numbers with the therapist.
Hence we decided to turn the CoWriter activity to teach numbers to the robot. Build on
existing iPad applications actually used by occupational therapists (Dawson Toth's \emph{ABC's
Writer}), we designed an Android application for pre-test and post-test. It consisted in drawing
numbers following an helping pattern, that becomes more fine with levels in
order to increase difficulty~\ref{fig:abc-writer}.\\

Since Henry was frequently drawing horizontally inverted numbers, or even
absolutely unrecognizable shapes, we programmed the robot to refuse shapes that
was too distant to a reference with a threshold we arbitrary fixed. In that way,
the child was forced to take care on what he was providing to the robot as
demonstration. Also, to make sure the robot was going to improve its handwriting
and to clearly show this improvement, we decided to make it absolutely starting
from scratch : for all numbers, the first try of the robot resulted in
a simple vertical stroke (see the first robot's try in~\ref{learning_6_demos}).

@@

\subsubsection{Results}
Despite his inattention, Henry was able to keep on the activity during more than
forty minutes in each session.\\

As soon as Henry understood that the robot was only accepting recognizable
shapes, he started to take care on the shapes he was providing to it. He used the
clearing button to retry when he judged that his numbers where not going to be
accepted by the robot. According to the therapist, it was the first time Henry
was intrinsically correcting himself.\\

Since the robot was learning from scratch, each time Henry succeeded to make his
demonstrations accepted by it, the improvement of its handwriting was
obvious~\ref{learning_6_demos},~\ref{henry_distances}.
Therefor, Henry quickly enter in an self-rewarding game where the rewards were
the improvement of the robot. Over the three last sessions he provided 99
demonstrations of number shapes to the robot.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_demos}
    \caption{\small Henry's demonstrations for the number ``6'' (top row) and
        corresponding shapes generated by the robot. After eight demonstrations,
        Henry decided that the robot's ``6'' was good enough, and went to
    another character: in that respect, he was the one leading the learning
process of the robot.}
    \label{learning_6_demos}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_distances}
    \caption{\small Two metrics to assess the handwriting progresses: Euclidian
    distance in the eigenspace of the number dataset (top figure) or in
cartesian space (bottom figure). Green lines represent the robot performance,
blue lines Henry's performance. The round IDs correspond to the demonstrations
pictured on Figure~\ref{henry_demos}.}
    \label{henry_distances}
\end{figure}

@@

The importance of the adult is further supported by our experiments: even with
an autonomous, nominally performing robot, to put the teacher's clothes on
remains (expectedly) difficult for 5-6 years old children, and during the
experiments we conducted, the adult always played a key role at prompting the
child to give feedback to the robot or to go to the next letter or word.  At a
higher orchestration level (and as reported in the two case studies with Diego
and Henry), the educational scenarios were also always designed and monitored by
the adults.

>>>>

Thomas is five and an half, and has been diagnosed with visuo-constructive
deficits. He is under the care of an occupational therapist, and tries to
workaround his inability to draw letters in a consistent manner. Vincent is six
and struggles at school with his poor handwriting and even poorer
self-confidence.

While Thomas is lively and always quick at shifting his attention from one
activity to another, Vincent is shy and poised. Two very different children,
facing however the same difficulty to write in a legible manner. And, hidden
beyond this impaired skill, psycho-social difficulties also arise: they
underperform at school, Thomas has to go for follow-up visits every week,
they both live under the label ``requires special care''. This is a source of
anxiety for the children and for their parents alike.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{henry}
    \caption{\small Thomas teaching Nao how to write numbers, with the help of an
    occupational therapist.}
    \label{fig:henry}
\end{figure}

@@

\begin{table}[ht!]
\centering
\caption{\small Field studies conducted within the project}
\label{studies}
\footnotesize
\begin{tabular}{@{}lp{4cm}p{2.2cm}p{1.8cm}l@{}}
\toprule
{\bf Study}     & {\bf Type}                                    & {\bf Duration}              & {\bf Children \#}                     & {\bf Ages} \\ \midrule
{\it ISG 1}     & Unstructured group interaction at school      & 16 min/group               & 4 $\times$ 8 children                 & 6-7        \\
{\it Florimont} & Individual/Pair interaction at school         & 11 min/group               & 7 (individual) + 7 $\times$ 2 (pairs) & 7-8        \\
{\it Vincent}     & Case-study, spanning over 4 weeks              & 1.5 hours $\times$ 4 weeks & 1                                     & 5          \\
{\it Haut-Lac}  & Pair (with turn-taking) interaction at school & 26 min/group                            & 7 $\times$ 2                          & 5-6        \\
{\it ISG 2}     & Individual interaction at school              & 18 min                            & 6                                     & 5-6        \\
{\it Thomas}     & Case-study, spanning over 4 weeks              & 1h $\times$ 4 weeks        & 1                                     & 5          \\ \bottomrule
\end{tabular}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Case Study 1: Vincent}

@@

Vincent is a six years old child with troubles in handwriting and a poor
self-esteem. Our goal was to figure out an environment in
order to make Vincent sustaining engagement over four sessions of one hour, one
session per week. We decided to introduce an appealing scenario that justifyed to the child
the activity where a robot wants to learn handwriting. We used two Nao robots :
a blue one (called Mimi) and an orange one (called Clem). Mimi was away for a
mission,and the two robots had to communicate by mails. But they decided to do
it ``like humans", with handwritten messages. While Mimi was a good in
handwriting, Clem had strong difficulties and needed the help of Vincent.\\

During the first session, Vincent just teached single letters to the robot.
To chose a letter, he had to pick it in a set of magneted plastic letters, and to
show it to the robot. But the robot was not programmed to detect and recognize
it. A third person was entering the letter and was sending it to the
system with the computer. With following sessions we gradually increased the
difficulty : in the second session the child teached full words. To chose the
word, he had to assemble plastic letters to build it and to show it to the
robot. In the two last sessions, he teached long sentences. This time he just
had to say the sentence to the robot, and then the third person was using the
compurter to enter it. \\

We designed a special scenario for the last session : Mimi came back to the
laboratory. In front of Vincent, Mimi said to Clem that it did not believe 
it realy improved its handwritting, and that the responses it recieved
was written by Vincent. To prove that it actualy improved its writing,
it gave to Clem thirty minutes for a final training. After the training, it
asked to Clem to write some words. Finally, he admitted that Clem actually
improved its handwriting.

\begin{figure}
    \centering
    \includegraphics[width=0.5\linewidth]{diego}
    \caption{\small Vincent working with Nao}
    \label{fig:diego}
\end{figure}

\subsubsection{Results}
In all sessions, Vincent was focused on the activity during forty minutes up to
one hour. In total, he provided 154 letters as demonstration.\\

Vincent accepted our scenario and seriously played the game. At the
beginning he seemed not confident and intimidated.  But quickly, he gain in
confidence to play with the robot. He even showed real affective bonds : he
sent an handwritten mail to it some days after the last session.\\

As a technical result, we do not know if Vincent improved his handwriting but he
likely conviced himself he was good enough at writing to help someone else (the
robot in that case). To show to Vincent how much he improved the robot's
handwriting, we printed the letter that the robot wrote during the third session
before and after his help (Fig.~\ref{fig:stimuli}).

\begin{figure}
    \centering
    \subfigure[Initial letter, generated by the robot]{
        \includegraphics[height=6cm]{diego-initial-letter}
    }
    \subfigure[Final letter, after training with Vincent]{
        \includegraphics[height=6cm]{diego-final-letter}
    }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Case Study 2: Thomas}

\subsubsection{Context, Study Design}
Thomas is a five and half years old child. It is followed by an occupational
therapist. According to the therapist, he presents visuo-constructive deficits.
From our perspective, the main effect of this trouble during handwriting
activity was that Thomas
did not have any mental template to draw the shape of a letter. So he was 
apparently performing random attempts and then was comparing with the provided template.
What is more, Thomas is strongly careless : he rarely payed attention to
advices, even to what he was doing when he was currently drawing, and he was
quickly shifting his attention from one
activity to another.
Hence, the goal of this study is to face those two challenging issues
(visuo-constructive deficits and inattention) in order to keep Thomas focused
on the activity during forty-minutes interactions, and to make the robot
evidently learning from his demonstrations.\\

This experiment took place in the therapist's office. It was divided in four
sessions as well. This time, we assumed that a scenario like the one we used for Vincent was
no longer relevant with Thomas. We just introduced the robot and quickly
said that it was seeking help to train for a robot handwriting contest.\\

Before the experiment, Thomas was working on writing numbers with the therapist.
Hence we decided to turn the CoWriter activity to teach numbers to the robot. Build on
existing iPad applications actually used by occupational therapists (Dawson Toth's \emph{ABC's
Writer}), we designed an Android application for pre-test and post-test. It consisted in drawing
numbers following an helping pattern, that becomes more fine with levels in
order to increase difficulty~\ref{fig:abc-writer}.\\

Since Thomas was frequently drawing horizontally inverted numbers, or even
absolutely unrecognizable shapes, we programmed the robot to refuse shapes that
was too distant to a reference with a threshold we arbitrary fixed. In that way,
the child was forced to take care on what he was providing to the robot as
demonstration. Also, to make sure the robot was going to improve its handwriting
and to clearly show this improvement, we decided to make it absolutely starting
from scratch : for all numbers, the first try of the robot resulted in
a simple vertical stroke (see the first robot's try in~\ref{learning_6_demos}).

@@

\subsubsection{Results}
Despite his inattention, Thomas was able to keep on the activity during more than
forty minutes in each session.\\

As soon as Thomas understood that the robot was only accepting recognizable
shapes, he started to take care on the shapes he was providing to it. He used the
clearing button to retry when he judged that his numbers where not going to be
accepted by the robot. According to the therapist, it was the first time Thomas
was intrinsically correcting himself.\\

Since the robot was learning from scratch, each time Thomas succeeded to make his
demonstrations accepted by it, the improvement of its handwriting was
obvious~\ref{learning_6_demos},~\ref{henry_distances}.
Therefor, Thomas quickly enter in an self-rewarding game where the rewards were
the improvement of the robot. Over the three last sessions he provided 99
demonstrations of number shapes to the robot.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_demos}
    \caption{\small Thomas's demonstrations for the number ``6'' (top row) and
        corresponding shapes generated by the robot. After eight demonstrations,
        Thomas decided that the robot's ``6'' was good enough, and went to
    another character: in that respect, he was the one leading the learning
process of the robot.}
    \label{learning_6_demos}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_distances}
    \caption{\small Two metrics to assess the handwriting progresses: Euclidian
    distance in the eigenspace of the number dataset (top figure) or in
cartesian space (bottom figure). Green lines represent the robot performance,
blue lines Thomas's performance. The round IDs correspond to the demonstrations
pictured on Figure~\ref{henry_demos}.}
    \label{henry_distances}
\end{figure}

@@

The importance of the adult is further supported by our experiments: even with
an autonomous, nominally performing robot, to put the teacher's clothes on
remains (expectedly) difficult for 5-6 years old children, and during the
experiments we conducted, the adult always played a key role at prompting the
child to give feedback to the robot or to go to the next letter or word.  At a
higher orchestration level (and as reported in the two case studies with Vincent
and Thomas), the educational scenarios were also always designed and monitored by
the adults.
